---
title: "Segmentation of EU countries, based on consumer & business indicators"
description: "[Oct 27, 2018] Segmentation of EU countries, including cluster analysis based on consumer & business confidence indicators"
output: html_document
---



In this article i used public Eurostat datasets, in order to develop a 
segmentation of the EU countries. In particular, the datasets used are:  
- **Consumer Confidence Indicator**  
- **Business 1**  
- **Business 1**  
- **Business 1**  
- **Business 1**  
- **Business 1**  

Explain more about the datasets, information, links, copy-paste from quides/site
SSSSSSSSS SSSSSSSSS  
fdfd
eeeeeeee
eeeeeeeeeeeeeee
eeeeeeeeeeeeeeeeeeeeeeeeee

The eurozone consists of 19 countries: Austria, Belgium, Cyprus, 
Estonia, Finland, France, Germany, Greece, Ireland, Italy, Latvia, Lithuania, 
Luxembourg, Malta, the Netherlands, Portugal, Slovakia, Slovenia, and Spain. 

The [Eurostat package](https://cran.r-project.org/web/packages/eurostat/index.html) 
used to obtain the original datasets.

More details about the ETL steps can be found, in the actual code, at the link 
at the end of the article.

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# Libraries
library(tidyverse)
library(rvest)
library(eurostat)


```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# Eurostat Business-Sentiment data ##############################################
dat <- get_eurostat("ei_bssi_m_r2", time_format = "date")
dat <- label_eurostat(dat)

sent <- 
  dat %>% 
  filter(s_adj == "Unadjusted data (i.e. neither seasonally adjusted nor calendar adjusted data)") %>% 
  select(geo, time, indic, values) %>% 
  spread(indic, values) %>% 
  rename(date = time) 

# Save the data
saveRDS(sent, file = "data/tidy/sent.RDS")
```



```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# Load the dataset
sent <- readRDS("/Users/manos/OneDrive/Projects/R/All_Projects/2018_10_ClusterAnalysis/data/tidy/sent.RDS")


# Only Consumer & business indicators
final <-
  sent %>%
  filter(date >= '2014-01-01') %>%
  select(-date) %>%
  group_by(geo) %>%
  summarise_all(median, na.rm = TRUE) %>% 
  mutate(Business = `Construction confidence indicator` + `Industrial confidence indicator` +
           `Retail confidence indicator` + `Services Confidence Indicator`,
         Consumers = `Consumer confidence indicator`)


```


# Exploratory Analysis

It is important here to check  

Below there is a scatterplot with marker labels indicating the positioning of 
each country in respect to consumer and confidence indicators.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

ggplot(final) + 
  geom_label(aes(Consumers, Business, label = geo))+
  labs(x = "Consumer Confidence",
       y = "Business Confidence",
       title = "scatterplot with marker labels")


ggplot(final, aes(Consumers, Business)) +
  geom_point()+
  geom_text(aes(label = geo), check_overlap = TRUE)+
  labs(x = "Consumer Confidence",
       y = "Business Confidence",
       title = "scatterplot with marker labels")

```

We can already indicate some patterns from the plot above:   
- There are some outliers, such as Greece(bottom left) & Sweden(top right)  
- There is a group of countries that are placed in the middle of the plot, indicating
average consumer & business confidence 
- There are a lot of countries that business confidence is substantially less than
consumer confidence

* The other way around doesn't happens



Country plot with business indicator
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}


```

* In the plot above we can see......


Country plot with confidence indicator
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}


```

* comments 

???? Maybe a Joy plot
```{r eval=FALSE, include=FALSE}

select(client_total_soft1, -exclude) %>% 
  mutate_all(scale) %>% 
  gather(key = "variable", value = "value") %>% 
  mutate(value = scale(value)) %>% 
  ggplot(aes(x = value,
             y = variable, fill = ..x..)) +
  ggridges::geom_density_ridges_gradient(rel_min_height = 0.0) +
  viridis::scale_fill_viridis(name = "") +
  ggridges::theme_ridges(font_size = 13, grid = TRUE) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 12),
        text = element_text(family = "mono")) +
  scale_x_continuous(limits = c(-2, 2))+
  labs(title = 'Density plots of variables',
       subtitle = 'All values are scaled')
```
* comments

# Modelling

## Hierarchical clustering

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(lubridate)

hier <- as.data.frame(final)
rownames(hier) <- hier$geo

hier <- dist(scale(hier[, -1]))
hier.cluster <- hclust(hier)


library(dendextend)
dend_test <- as.dendrogram(hier.cluster)
plot(dend_test, main = "Hierarchical clustering")


dend_2 <- color_branches(dend_test, h = 6)
plot(dend_2, main = "Hierarchical clustering & cut-off value = 35 ")

```



## K-means Check optimal number of k with elbow (scree) plot

Total within cluster sum of squares (the sum of euclidean distances between each
observation and the centroid corresponding to the cluster to which the observation 
is assigned)

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(purrr)
library(cluster)


kmeans.dat <- dist(scale(final[, 2:6]))

# Use map_dbl to run many models with varying value of k (centers)
tot_withinss <- map_dbl(1:10,  function(k){
  model <- kmeans(x = kmeans.dat, centers = k)
  model$tot.withinss
})

# Generate a data frame containing both k and tot_withinss
elbow_df <- data.frame(
  k = 1:10,
  tot_withinss = tot_withinss
)


ggplot(elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:10) +
  ylab("Total within sum of squares")

```


## K-means (All variables) silhouette analysis 

It determines how well each of the observations fit into corresponding cluster (Higher is better).
It involves calculating a measurement called the silhouette width for every 
observation which consists of 2 parts i) the within cluster distance C (average 
distance from that observation to every other observation within the cluster) and
ii) the closest neighbor distance N (the average distance from that observation 
to the points of the closest neighboring cluster).  
Calculation:  
- if C < N then 1 - C/N  
- if C = N then 0  
- if C > N then N/C - 1

Interpretation:
- A value close to 1 suggests that this observation is well matched to its current cluster.  
- A value of 0 suggests that it is on the border between two clusters and can possibly 
belong to either one.  
- A value close to -1 suggests that the observation has a better fit to its 
closest neighboring cluster.



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(purrr)
library(cluster)
# Use map_dbl to run many models with varying value of k
sil_width <- map_dbl(2:10,  function(k){
  model <- pam(x = kmeans.dat, k = k)
  model$silinfo$avg.width
})

# Generate a data frame containing both k and sil_width
sil_df <- data.frame(
  k = 2:10,
  sil_width = sil_width
)

# Plot the relationship between k and sil_width
ggplot(sil_df, aes(x = k, y = sil_width)) +
  geom_line() +
  scale_x_continuous(breaks = 2:10)+
  ylab("Average Silhouette width")

```


## K-means (All variables) explore variables

**Below there is a table with information about all clusters (with all variables)**


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
set.seed(42)

# Build a k-means model for the customers_spend with a k of 2
model_customers <- kmeans(kmeans.dat, centers = 4)

# Extract the vector of cluster assignments from the model
clust_countries <- model_customers$cluster

# Build the segment_customers dataframe
segment_countries <- mutate(final[1:6], cluster = clust_countries)

library(DT)
# Calculate the mean for each category
segment_countries %>% 
  group_by(cluster) %>% 
  # summarise_all(funs(round(mean(.),2))) %>% 
  # datatable(filter = 'top', options = list(autoWidth = TRUE))
  add_tally() %>% 
  summarise_each(funs(round(mean(., na.rm = TRUE),2))) %>% 
  select(1, "n", 3:ncol(final)) %>% 
  datatable(filter = 'top', options = list(pageLength = 5, autoWidth = TRUE, dom = 'pt'))


```


**Below there is a table of all partners indicating their cluster (All variables) **

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(DT)
# Print the table with 
segment_countries %>% 
  select(c(1, "cluster", 2:6)) %>% 
  datatable(filter = 'top', options = list(pageLength = 10, autoWidth = TRUE, dom = 'pt'))

```



## Map countries with cluster

```{r}

```







